---
title: "Trang Quach - TidyTuesday I"
output: html_document
date: '10/04/2019'
editor_options: 
  chunk_output_type: console
---
## Data loading
```{r, echo=TRUE, message=FALSE}
pizza_jared <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_jared.csv")
pizza_barstool <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_barstool.csv")
pizza_datafiniti <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2019/2019-10-01/pizza_datafiniti.csv")

```

## Loading package

```{r, echo=TRUE, message=FALSE}
library(tidyverse)
library(zipcode)
library(DT)
data(zipcode) # I used this data to add the state information to pizza_barstool which only has zipcode
```

## Using pizza_datafiniti to map 9 most frequent brand names of Pizza
### Cleaning data
```{r}
location <- pizza_datafiniti %>%
  #remove the duplicates in dataset
  distinct(longitude, latitude, .keep_all = TRUE) %>%
  # Not look at AK or HI state
  filter(!province %in% c('AK', 'HI')) %>% 
  #Correct some restaurant name
  mutate(name = str_remove(name, "[']")) %>%
  mutate(name = replace(name, name=="Papa Murphys Pizza", "Papa Murphy's"))

```
### Each restaurant is present at how many states
```{r}
number_state <- location %>%
  group_by(name, province) %>%
  summarise(number = n()) %>%
  select(-number) %>%
  summarise(number_state= n()) %>%
  arrange(desc(number_state)) 

datatable(number_state%>% top_n(15, number_state), options = list(pageLength=10))
```

* Papa John Pizza are most popular, which is present at 24 states. 

* Little Caesars Pizza is present at 20 states.

* Only 5 brand names are present at more than 4 states.

```{r}
# calculcate the total number of restaurants for a specific name
namesum <- location %>%
  group_by(name) %>%
  summarise(total = n()) %>% #total is total number of restaurants with a specific name 
  arrange(desc(total))

datatable(namesum %>% top_n(25, total), options = list(pageLength=10))

```

* Papa John Pizza has the highest number of stores in US with 68 restaurants, following by Little Caesars Pizza with 38. 

* Only 8 pizza brands has more than 9 restaurants in US.


### Merge 3 datasets. This data is used for mapping
```{r}
location1 <- location %>% 
  left_join(number_state, by="name") %>%
  left_join(namesum, by="name") %>%
  mutate(price_max = if_else(price_range_max %in% c(20,25), "20-25", "40"))
```


### Number of restaurants in a specific state
```{r}
location1 %>%
  count(province, sort=TRUE)
```

* The states having the highest number of pizza restaurants is New York with 424. 

* The following is California with 255. 

* Only 6 states have more than 100 Pizza restaurants.

## Top 9 restaurants with the highest numbers and how these restaurants distributed in US
```{r}

location1 %>% 
  filter(as.integer(fct_reorder(name, total, .desc = TRUE)) <= 9) %>%
  mutate(name = glue::glue("{name} ({total})"), 
         name = fct_reorder(name, desc(total))) %>%
  ggplot(aes(longitude, latitude, color=price_max))+
  borders("state") +
  geom_point(alpha=0.5) +
  coord_map()+
  facet_wrap( ~ name) +
  theme_void() +
  labs(title = "How pizza restaurant distributed in US and how max price changes by distribution",
       subtitle = "Only 9 restaurants with the highest numbers were presented. Number of restaurants in parentheses.")
```

* Papa Johns Pizza and Little Caesars Pizza, and Sbarro spreaded out in almost regions in US, except for parts of west and midwest.

* Papa Murphys did not cover the Southeast and Northeast areas. In reverse, Pizza Hut and Jets Pizza covered Southeast and Northeast, but did not cover the remainder area.

* Barros Pizza was only present at Arizona.

* With the same brand, the maximum price of Hungry Howies Pizza and Barros Pizza varied by locations.


##  Rating of these 9 restaurant in terms of price and quality
```{r}
x1 <- location1 %>% 
  filter(as.integer(fct_reorder(name, total, .desc = TRUE)) <= 9) %>% distinct(name)
x2 <- pizza_jared %>% distinct(place)
sum(x2 %in% x1) # All of 9 restaurants are not in pizza_jared


x3 <- pizza_barstool %>% distinct(name)
sum(x3 %in% x1) # All of 9 restaurants are not in pizza_barstool
```
* I cannot evaluate the price and quality of the 9 restaurants because pizza_barstool and pizza_jared did not include the information for these restaurants.

## Using data pizza_bartool
### Map to see what states has highest number of pizza restaurants and how price they are
```{r}
zipcode <- zipcode %>% select(zip, state) %>% mutate(zip = as.numeric(zip))
data2 <-pizza_barstool %>%
  distinct(longitude, latitude, .keep_all = TRUE) %>%
  filter(!is.na(longitude) & !is.na(latitude)) %>%
  left_join(zipcode, by="zip") %>%
  group_by(state) %>%
  summarise(total= n(), long = mean(longitude), lat =mean(latitude), 
            mean_price= mean(price_level))

data2 %>%
  ggplot(aes(long, lat, size=total, color= mean_price))+
  geom_point()+
  borders("state")+
  coord_map() +
  theme_void()
```

* In the shared boder of New Jersey and New York, there were more than 200 pizza restaurants. 

* In Massachusetts, there were more than 50 pizza restaurants. The price of pizza in these places was cheap in average.


## How scores varied according to providers' rating and price
```{r}
a1 <- pizza_barstool %>% 
  gather(key="type", value ="score", 11,14,17,20) %>%
  select(-c(11,12,13,14,15,16,17,18))

a1 %>%
  ggplot(aes(factor(provider_rating), score))+
  geom_boxplot()+
  geom_jitter(alpha=0.1, width=0.3)+
  facet_wrap(~type)

a1 %>% 
  ggplot(aes(factor(price_level), score))+
  geom_boxplot()+
  geom_jitter(alpha=0.1, width=0.3)+
  facet_wrap(~type)

```

* Except for critic score, all scores seemed to increase when provider rating increased. Most of rating were for category of 3, 3.5, and 4 of provider_rating

* Similarly, all scores seemed to increase when price level increased, except for critic score. Most of rating were for category 1 and 2 of price level.

* For critic score, regardless of the provider_rating or price level, this score mainly graded as 0.